---
title: "May Tabular Playground"
---

```{r}
library(tidyverse)
library(eemisc)
library(janitor)
library(corrplot)
library(psych)

theme_set(theme_minimal())

trn <- read_csv(here::here("may_22_tabular/data/train.csv"))
tst <- read_csv(here::here("may_22_tabular/data/test.csv"))

```

# Exploring the Target

```{r}
mean(trn$target)
```

# Explore Predictors

```{r}
trn |>
    select(where(is.numeric)) |>
    summary()
```

ok, so this is actually some combination of integer data (maybe ordinal; e.g. `f_30`), numeric data (e.g. `f_23`), and binary data (e.g. `f_29`). It's also worth noting that som e of the numeric data is on very different scales, so we'll need to normalize the data.

Let's check out the correlations among the numeric variables next:

```{r}
trn |>
    select(where(is.numeric), -id) |>
    cor() |>
    corrplot()
```

So the correlations here are mostly pretty small. `f_28` appears to be moderately correlated with a few other variables, but there's not really much going on here. And there really isn't much that's correlated with the target, either.

Let's use `{skimr}` to get a better picture of the data here

```{r}
trn |>
    select(where(is.numeric), -id) |>
    skim()
```

The distributions are really helpful here. Basically, what we can see is:

- `f_00` through `f_06` are normally distributed with mean 0 and sd 1;
- `f_07` through `f_18` roughly follow a Poisson distribution or a negative-binomial distribution (maybe?). In other words, they appear to be counts.
- `f_19` through `f_28` also appear to be normally distributed, although maybe slightly skewed or slightly off-center.
- `f_29` and `f_30` are counts with few unique values. `f_29` is binary, and `f_30` has 3 values.

There's also that one string predictor that we'll want to play around with. Let's look at how many unique values there are for it -- I'm assuming a bunch:

```{r}
length(unique(trn$f_27))
```

And what proportion of the rows contain a unique value here?

```{r}
length(unique(trn$f_27))/nrow(trn)
```

Yikes. 82% are unique values. It might be worth thinking about this almost like PCA -- think of the first letter like the first component, the second letter like the seonc component, etc., and then use only like the first three characters in the string to put data into groups (could use more; 3 is just like a heuristic).