---
title: "EDA"
author: "EE"
---
    
# EDA of March 2022 Tabular Playground Data

```{r}
library(tidyverse)
library(eemisc)
library(lubridate)
library(hms)

trn <- read_csv(here::here("mar_22_tabular/data/train.csv"))

theme_set(theme_ee(accent_color = "gray80"))
```


```{r}
dim(trn)
```


```{r}
head(trn)
```

# Missingness

```{r}
summarize(trn, across(everything(), ~ sum(is.na(.x))/length(.x)))
```

No missing data across any of our cols

# Target: Congestion

```{r}
ggplot(trn, aes(x = congestion)) +
    geom_histogram()
```

Sort of normally distributed, with some spikes at certain values. Like the directions for the competition say, this has been scaled so measurements are 0-100.

```{r}
c(mean(trn$congestion), sd(trn$congestion))
```

We can also see the number of unique values:
```{r}
length(unique(trn$congestion))
```

So, every value from 0-100 is represented, which is useful to know when predicting values.

# Continuous Predictors (but not time)

```{r}
summary(cbind(trn$x, trn$y))
```

So these really don't seem like continuous predictors? I think they're more like ordinal values?

```{r}
trn %>%
    select(x, y) %>%
    pivot_longer(cols = c("x", "y"), names_to = "key", values_to = "val") %>%
    count(key, val) %>%
    ggplot(aes(x = n, y = as.character(val))) +
    geom_col() +
    facet_wrap(~key)
```

Let's see if there's any difference in the distriubtion of the outcome by x and y

```{r}
trn %>%
    mutate(across(c("x", "y"), as.character)) %>%
    ggplot(aes(x = congestion)) +
    geom_histogram() +
    facet_grid(rows = vars(x), cols = vars(y))
```

So, there are some differences here.

# Direction Predictors 

```{r}
trn %>%
    count(direction, sort = TRUE)
```

And let's do the same thing above & look at distribution differences between these:

```{r}
trn %>%
    ggplot(aes(x = congestion)) +
    geom_histogram() +
    facet_wrap(vars(direction))
```

# Date Time Predictors

So, this is going to be the money predictor here.
```{r}
trn %>%
    mutate(
        tm = as_hms(time),
        wkday = wday(time, label = TRUE),
        mo = month(time, label = TRUE)
    ) %>%
    glimpse()
    
```