---
title: "House Price Mod"
author: "EE"
date: "4/25/2020"
output: html_document
---

The purpose of this notebook is to show how to do some basic feature engineering and then fit an xgboost model using the tidymodels framework.
```{r setup}

library(tidyverse)
library(tidymodels)
library(janitor)
library(skimr)

theme_set(theme_minimal())

train <- read_csv(here::here("House Prices/Data/train.csv")) %>%
  clean_names()

test <- read_csv(here::here("House Prices/Data/test.csv")) %>%
  clean_names()

miss_plot <- function(data, color1 = "steelblue1", color2 = "steelblue4", bound = 0) {
  miss_tab <<- tibble(
    column = names(data),
    perc_miss = map_dbl(data, function(x) sum(is.na(x))/length(x))
  ) %>%
    filter(perc_miss > bound)
  
  ggplot(miss_tab, aes(x = column, y = perc_miss)) +
    geom_bar(stat = "identity", aes(fill = ..y..)) +
    scale_y_continuous(labels = scales::percent) +
    theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
    scale_fill_gradient(low = color1, high = color2, name = "Percent Missing") +
    labs(
      title = "Missingness by Variable",
      y = "Percent Missing",
      x = "Variables"
    )
}
```


First, let's take a quick look at our training data using the skim() function from the skimr package.
```{r skim}
skim(train)
```

Let's look at the log of the lot area, since the raw lot area is heavily skewed.
```{r}
ggplot(train) +
  geom_histogram(aes(log(lot_area)))
```

We'll probably want to log every area/square footage variable in the dataset, but let's take a look at a couple more.
```{r}
ggplot(train) +
  geom_histogram(aes(log(x1st_flr_sf)))

ggplot(train) +
  geom_histogram(aes(log(x2nd_flr_sf)))
```

And let's also look at missingness across variables. 
```{r}
miss_plot(train)
```

Some of the missingness here is actually meaningful -- i.e. missing data on the "alley" variable means that the house doesn't have alley access, etc. The easiest approach for dealing with this in nominal variables will just be to dummy out the factors and let NA be its own level. For numeric variables, I'll impute using knn.

## Preprocessing Recipe

This step sets up a recipe that will create the design matrix for our model. To learn more about the preprocessing steps, check out the recipes package.
```{r recipe}
pre_rec <- recipe(sale_price ~ ., data = train) %>%
  update_role(id, new_role = "ID") %>%
  step_string2factor(all_nominal()) %>%
  step_other(all_nominal(), threshold = .05) %>%
  step_log(matches("sf|area|lot_frontage"), offset = 1) %>%
  step_log(all_outcomes(), skip = TRUE) %>%
  step_knnimpute(all_numeric(), -all_outcomes(), neighbors = 5) %>%
  step_center(all_numeric(), -all_outcomes()) %>%
  step_scale(all_numeric(), -all_outcomes()) %>%
  step_dummy(all_nominal()) %>%
  step_nzv(all_predictors())
```


## Setting up Resamples
Bootstrapping 25 resamples here to train on.
```{r boots}
set.seed(0408)
train_boot <- bootstraps(train, times = 25)
```

## Setting up the Model

Here, I'm setting up an xgboost model with a hardcoded min_n, number of trees, and learning rate. I'll tune the values for mtry and tree depth during training.
```{r}
gbm_mod <- boost_tree() %>%
  set_args(
    mtry = tune(),
    min_n = 10,
    tree_depth = tune(),
    trees = 2000,
    learn_rate = .01,
  ) %>%
  set_engine("xgboost") %>%
  set_mode("regression")
```

## Putting Model and Recipe into a Workflow

This step puts the previous recipe and the model we specified into a workflow object, which makes it easier to carry these around for later fitting.
```{r workflow}
gbm_work <- workflow() %>%
  add_recipe(pre_rec) %>%
  add_model(gbm_mod) 
```


## Setting up Tuning Grid & Fitting

Here I'm setting up values for mtry to tune on. I'm using the default values for tree_depth. Afterward, I'm specifying a grid of values to tune across and then fitting the model.
```{r}
gbm_params <- parameters(gbm_work)

gbm_params <- gbm_params %>%
  update(mtry = mtry(c(1L, 10L)))

gbm_grid <- gbm_params %>%
  grid_max_entropy(size = 20)

doParallel::registerDoParallel()
set.seed(0408)
gbm_tuned <- gbm_work %>%
  tune_grid(
    resamples = train_boot,
    grid = gbm_grid,
    metrics = metric_set(rmse)
  )
```

## Selecting the Best Parameters & Finalizing Model

Here, I'm choosing the best model parameters from the fitting, finalizing the workflow to incorporate these optimal parameters, and fitting the final model.
```{r finalize model}
best_params <- select_best(gbm_tuned, metric = "rmse", maximize = FALSE)

best_params

gbm_work <- gbm_work %>%
  finalize_workflow(best_params)

gbm_fit <- fit(gbm_work, data = train)
```

## Predicting Test Data
```{r predict}
gbm_preds <- predict(gbm_fit, new_data = test)

exp_preds <- exp(gbm_preds) %>% as_vector()

gbm_sub <- tibble(
  id = test$id,
  SalePrice = exp_preds
)

write_csv(gbm_sub, "submission.csv")
```

