---
title: "House Price Mod"
author: "EE"
date: "4/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidymodels)
library(janitor)
library(skimr)

theme_set(theme_minimal())

train <- read_csv(here::here("House Prices/Data/train.csv")) %>%
  clean_names()

test <- read_csv(here::here("House Prices/Data/test.csv")) %>%
  clean_names()

miss_plot <- function(data, color1 = "steelblue1", color2 = "steelblue4", bound = 0) {
  miss_tab <<- tibble(
    column = names(data),
    perc_miss = map_dbl(data, function(x) sum(is.na(x))/length(x))
  ) %>%
    filter(perc_miss > bound)
  
  ggplot(miss_tab, aes(x = column, y = perc_miss)) +
    geom_bar(stat = "identity", aes(fill = ..y..)) +
    scale_y_continuous(labels = scales::percent) +
    theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
    scale_fill_gradient(low = color1, high = color2, name = "Percent Missing") +
    labs(
      title = "Missingness by Variable",
      y = "Percent Missing",
      x = "Variables"
    )
}
```

```{r}
skim(train)
```

Let's look at the log of the lot area, since the raw lot area is heavily skewed.
```{r}
ggplot(train) +
  geom_histogram(aes(log(lot_area)))
```

We'll probably want to log every area/square footage variable in the dataset, but let's take a look at a couple more.
```{r}
ggplot(train) +
  geom_histogram(aes(log(x1st_flr_sf)))

ggplot(train) +
  geom_histogram(aes(log(x2nd_flr_sf)))
```

And let's also look at missingness
```{r}
miss_plot(train)
```

We'll want to come up with an approach for dealing with these. With a note that missingness on some of these is actually an indicator that the feature is absent/doesn't exist (e.g. fence). See below for cleaning steps.

## Preprocessing Recipe
```{r recipe}
pre_rec <- recipe(sale_price ~ ., data = train) %>%
  update_role(id, new_role = "ID") %>%
  step_string2factor(all_nominal()) %>%
  step_other(all_nominal(), threshold = .05) %>%
  step_log(matches("sf|area|lot_frontage"), offset = 1) %>%
  step_log(all_outcomes(), skip = TRUE) %>%
  step_knnimpute(all_numeric(), -all_outcomes(), neighbors = 5) %>%
  step_center(all_numeric(), -all_outcomes()) %>%
  step_scale(all_numeric(), -all_outcomes()) %>%
  step_dummy(all_nominal()) %>%
  step_nzv(all_predictors())
```


## Setting up Resamples
Bootstrapping 25 resamples here to train on
```{r boots}
set.seed(0408)
train_boot <- bootstraps(train, times = 25)
```

## Setting up the Model
```{r}
gbm_mod <- boost_tree() %>%
  set_args(
    mtry = tune(),
    min_n = 10,
    tree_depth = tune(),
    trees = 2000,
    learn_rate = .01,
  ) %>%
  set_engine("xgboost") %>%
  set_mode("regression")
```

## Putting Model and Recipe into a Workflow
```{r workflow}
gbm_work <- workflow() %>%
  add_recipe(pre_rec) %>%
  add_model(gbm_mod) 
```


## Setting up Tuning Grid & Fitting
```{r}
gbm_params <- parameters(gbm_work)

gbm_params <- gbm_params %>%
  update(mtry = mtry(c(1L, 10L)))

gbm_grid <- gbm_params %>%
  grid_max_entropy(size = 20)

doParallel::registerDoParallel()
set.seed(0408)
gbm_tuned <- gbm_work %>%
  tune_grid(
    resamples = train_boot,
    grid = gbm_grid,
    metrics = metric_set(rmse)
  )
```

## Selecting the Best Parameters & Finalizing Model
```{r finalize model}
best_params <- select_best(gbm_tuned, metric = "rmse", maximize = FALSE)

best_params

gbm_work <- gbm_work %>%
  finalize_workflow(best_params)

gbm_fit <- fit(gbm_work, data = train)
```

## Predicting Test Data
```{r predict}
gbm_preds <- predict(gbm_fit, new_data = test)

exp_preds <- exp(gbm_preds) %>% as_vector()

gbm_sub <- tibble(
  id = test$id,
  SalePrice = exp_preds
)

write_csv(gbm_sub, here::here("House Prices/Submissions/gbm_preds.csv"))
```

