# April 2022 Tabular Playground EDA 


```{r}
 library(tidyverse)
 library(janitor)
 library(eemisc)

 trn <- read_csv(here::here("apr_22_tabular/data/train.csv"))
 trn_labs <- read_csv(here::here("apr_22_tabular/data/train_labels.csv"))
 tst <- read_csv(here::here("apr_22_tabular/data/test.csv"))

 ## joining subject into trn_labs
 trn_labs <- trn_labs %>%
     left_join(trn %>% distinct(subject, sequence))

 theme_set(theme_ee(accent_color = "grey80"))
```

ok, so, steps within sequences within subjects. Labels are associated with each sequence, though, not each step.

# Target

```{r}
table(trn_labs$state)
```

So these states are pretty much equally likely, which is actually good -- we don't have to deal with class imbalance.

# Participants & Sequences

How many participants do we have?

```{r}
length(unique(trn$subject))
```

And how many unique sequences?

```{r}
length(unique(trn$sequence))
```

Let's look to see if sequences are evenly distributed across subjects:

```{r}
trn %>%
    distinct(subject, sequence) %>%
    count(subject) %>%
    ggplot(aes(x = n)) +
    geom_histogram()
```

Right -- so our distribution here is slightly skewed. Some subjects account for a lot of sequences, whereas others only account for a few. This will make nesting obs difficult to some extent, although there are def ways to handle it.

We saw previously that `state` is pretty balanced -- we saw about as many 0s as 1s across all sequences. But let's see if participants differ in their average states

```{r}
trn_labs %>%
    group_by(subject) %>%
    summarize(state = mean(state)) %>%
    ggplot(aes(x = state)) +
    geom_histogram()
```


Ok, so, yeah -- people are differing quite a bit in their average state. There's a big spike at 0 -- my assumption here is that people with relatively few trials tend to score 0, but let's look at that

```{r}
trn_labs %>%
    group_by(subject) %>%
    summarize(state = mean(state), n = n()) %>%
    select(-subject) %>%
    cor()
```

This is useful -- there's a pretty good sized correlation between number of trials (within a participant) and average state.

## RESUME HERE.

Thoughts:

- PCA on sensors might be useful
- LPA within each sequence on the PCs?
- Could also do mean/var of each sensor within each sequence
- note that n_trials is correlated with avg subject state at .68
