# April 2022 Tabular Playground EDA 


```{r}
 library(tidyverse)
 library(janitor)
 library(eemisc)
library(ggridges)

 trn <- read_csv(here::here("apr_22_tabular/data/train.csv"))
 trn_labs <- read_csv(here::here("apr_22_tabular/data/train_labels.csv"))
 tst <- read_csv(here::here("apr_22_tabular/data/test.csv"))

 ## joining subject into trn_labs
 trn_labs <- trn_labs %>%
     left_join(trn %>% distinct(subject, sequence))

 theme_set(theme_ee(accent_color = "grey80"))
```

ok, so, steps within sequences within subjects. Labels are associated with each sequence, though, not each step.

# Target

```{r}
table(trn_labs$state)
```

So these states are pretty much equally likely, which is actually good -- we don't have to deal with class imbalance.

# Participants & Sequences

How many participants do we have?

```{r}
length(unique(trn$subject))
```

And how many unique sequences?

```{r}
length(unique(trn$sequence))
```

Let's look to see if sequences are evenly distributed across subjects:

```{r}
trn %>%
    distinct(subject, sequence) %>%
    count(subject) %>%
    ggplot(aes(x = n)) +
    geom_histogram()
```

Right -- so our distribution here is slightly skewed. Some subjects account for a lot of sequences, whereas others only account for a few. This will make nesting obs difficult to some extent, although there are def ways to handle it.

We saw previously that `state` is pretty balanced -- we saw about as many 0s as 1s across all sequences. But let's see if participants differ in their average states

```{r}
trn_labs %>%
    group_by(subject) %>%
    summarize(state = mean(state)) %>%
    ggplot(aes(x = state)) +
    geom_histogram()
```


Ok, so, yeah -- people are differing quite a bit in their average state. There's a big spike at 0 -- my assumption here is that people with relatively few trials tend to score 0, but let's look at that

```{r}
trn_labs %>%
    group_by(subject) %>%
    summarize(state = mean(state), n = n()) %>%
    select(-subject) %>%
    cor()
```

This is useful -- there's a pretty good sized correlation between number of trials (within a participant) and average state.

# Sensor Data

Let's look at the bio sensor data. To start, let's summarize each sensor:

```{r}
sensor_cols <- names(trn[, -c(1:3)])

summary(trn[, sensor_cols])
```


So, some of these values are going to need to get cleaned. Let's take a closer look at where the majority of the data falls

```{r}
lapply(trn[, sensor_cols], quantile, probs = c(.05, .95))
```

Sensor 12 is still a bit wonky, but the rest seem to have most of their data within +/- 2. The strategy here might be to replace extreme values with the 5th & 95th percentiles, plus add a tiny bit of random noise to each?

*Note: it might be a better idea to just normalize these and not manually deal with the extreme values.*

```{r}
trn <- trn %>%
    mutate(across(starts_with("sensor"), function(x) {
        (x - mean(x)) / sd(x)
    }))
```
 
 *it might be worthwhile to drop extreme values from the training set later on*


Ok, so now let's look at the distribution of each 

```{r}
trn %>%
    pivot_longer(cols = starts_with("sensor"), names_to = "sensor", values_to = "val") %>%
    ggplot(aes(x = val, y = sensor)) +
    geom_density_ridges()
```

And let's do it by state:

```{r}
trn %>%
    pivot_longer(cols = starts_with("sensor"), names_to = "sensor", values_to = "val") %>%
    left_join(trn_labs[, c("sequence", "state")]) %>%
    ggplot(aes(x = val, y = sensor, fill = as.character(state))) +
    geom_density_ridges(alpha = .4)
```


alright, so there's basically no difference here between the distributions of state 0 and state 1 for any of these sensors. We still have an issue with extreme values, too. This might just be something to address when modeling.

Let's check out the correlations among sensors:

```{r}
cor(trn[, sensor_cols])
```

So, most of these are tiny, but there are a few that are moderate. That said, PCA may not do much for me here if the correlations are so small.

Let's get the median of each sensor per sequence and see the distribution there. We'll want to keep in mind that there are still person-level dependencies we're not capturing here, though:

```{r}
med_sensors <- trn %>%
    group_by(sequence) %>%
    summarize(across(starts_with("sensor"), median))

med_sensors %>%
    pivot_longer(cols = starts_with("sensor"), names_to = "sensor", values_to = "val") %>%
    left_join(trn_labs[, c("sequence", "state")]) %>%
    ggplot(aes(x = val, y = sensor, fill = as.character(state))) +
    geom_density_ridges(alpha = .4)
```


Still not a ton here that seems super informative, although sensor 2 has some strange multi-modal peak stuff going on.

Let's do the same thing with standard deviations, although I expect this to be kinda wonky due to some of the extreme values

```{r}
trn %>%
    group_by(sequence) %>%
    summarize(across(starts_with("sensor"), sd)) %>%
    pivot_longer(cols = starts_with("sensor"), names_to = "sensor", values_to = "val") %>%
    left_join(trn_labs[, c("sequence", "state")]) %>%
    ggplot(aes(x = val, y = sensor, fill = as.character(state))) +
    geom_density_ridges(alpha = .4)
```

Right, so still not a whole lot to go on here.

```{r}
#function to compute rci
std_err <- function(x) {
    sd(x) / sqrt(length(x))
}

#sort of a bastardized reliable change, but w/e
rci <- function(x1, x2) {
    se <- std_err(x1)

    sdiff <- sqrt(2 * se^2)

    v <- (x1 - x2) / sdiff

    abs(v) > 1.96
}
```

```{r}
rc_summed <- trn %>%
    pivot_longer(cols = starts_with("sensor"), names_to = "sensor", values_to = "val") %>%
    group_by(sequence, sensor) %>%
    arrange(sequence, sensor, step) %>%
    mutate(lag_val = lag(val), rci = rci(val, lag_val)) %>%
    summarize(prop_rc = sum(rci, na.rm = TRUE) / 59) %>%
    ungroup() %>%
    pivot_wider(names_from = sensor, values_from = prop_rc, names_glue = "{sensor}_prop_rc")
```

Thoughts:

- See this chapter from MK's feat eng book: https://bookdown.org/max/FES/profile-data.html
- PCA on sensors might be useful
- LPA within each sequence on the PCs?
- Could also do mean/var of each sensor within each sequence
- Number of reliable changes within a given sequence? Look more into reliable change indices
- note that n_trials is correlated with avg subject state at .68
