---
title: "Data modeling"
author: "EE"
date: "12/19/2019"
output: html_document
---

# Note to self -- if this doesn't work, try to read in the files without the event_data column to see if that helps

Ok, so this model is actually worse than the median only model :)

Something to learn from moving forward

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(tidymodels)

set.seed(0408)

#Reading in data and cleaning
train <- data.table::fread(here("2019 Data Science Bowl/Data/train.csv"))
test <- data.table::fread(here("2019 Data Science Bowl/Data/test.csv"))
trn_lbls <- data.table::fread(here("2019 Data Science Bowl/Data/train_labels.csv"))
specs <- data.table::fread(here("2019 Data Science Bowl/Data/specs.csv"))

#getting the last assessment in the test set
last_test_assessment <- test %>%
  filter(type == "Assessment") %>%
  arrange(desc(timestamp)) %>%
  distinct(installation_id, .keep_all = TRUE) %>%
  select(installation_id, game_session, title)

#creating a function to score events
score_events <- function(df){
  install_ids <- df %>% 
    distinct(game_session, installation_id, .keep_all = T) %>%
    select(game_session, installation_id, timestamp, title)
  
  df %>% 
    filter(str_detect(event_data, '""correct""')) %>%
    mutate(correct = as.numeric(str_detect(event_data, '""correct"":true'))) %>%
    group_by(game_session) %>%
    summarize(num_correct = sum(correct),
              num_incorrect = n() - sum(correct),
              accuracy = sum(correct) / n()) %>%
    select(game_session, num_correct, num_incorrect, accuracy) %>%
    left_join(., install_ids, by = c('game_session'))
}

```

## Getting performance history for training and merging with data
```{r}
train <- train %>%
  filter(type == "Game") %>%
  score_events() %>%
  mutate(game_title = title) %>%
  group_by(installation_id, game_title) %>%
  summarize(
    sum_correct = sum(num_correct),
    sum_incorrect = sum(num_incorrect),
    mean_accuracy = mean(accuracy, na.rm = TRUE),
    med_accuracy = median(accuracy, na.rm = TRUE),
    sd_accuracy = sd(accuracy, na.rm = TRUE),
    n_games = n(),
    n_distinct_games = n_distinct(game_title),
    med_mean_diff_acc = med_accuracy - mean_accuracy
  ) %>%
  ungroup()

train<- trn_lbls %>%
  left_join(train, by = "installation_id") %>%
  distinct(installation_id, game_session, game_title, .keep_all = TRUE) %>%
  pivot_wider(names_from = game_title,
              values_from = c(sum_correct, sum_incorrect, mean_accuracy, med_accuracy, sd_accuracy, n_games,
                              n_distinct_games, med_mean_diff_acc)) %>%
  select(-matches("_NA$")) %>%
  mutate_at(vars(matches("acc")),
            ~replace_na(., median(., na.rm = TRUE))) %>%
  mutate_at(vars(matches("games|sum")),
            ~replace_na(., 0)) %>%
  janitor::clean_names() %>%
  select(-c("game_session", "installation_id", "num_correct", "num_incorrect", "accuracy")) %>%
  mutate(accuracy_group = as.factor(accuracy_group))
```

Training a random forest model on the train data
```{r rf train}
#setting my preprocessing
pre_recipe <- train %>%
  recipe(accuracy_group ~ .) %>%
  step_center(all_numeric(), -all_outcomes()) %>%
  step_scale(all_numeric(), -all_outcomes()) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  prep()

#getting the dataframe to work with
train <- juice(pre_recipe)

#training a random forest model
rf_trn <- rand_forest(trees = 500, mode = "classification") %>%
  set_engine("ranger") %>%
  fit(accuracy_group ~ ., data = train)

rm(train)
```

## Getting performance history on test set
```{r test hist}
test <- test %>%
  filter(type == "Game") %>%
  score_events() %>%
  mutate(game_title = title) %>%
  group_by(installation_id, game_title) %>%
  summarize(
    sum_correct = sum(num_correct),
    sum_incorrect = sum(num_incorrect),
    mean_accuracy = mean(accuracy, na.rm = TRUE),
    med_accuracy = median(accuracy, na.rm = TRUE),
    sd_accuracy = sd(accuracy, na.rm = TRUE),
    n_games = n(),
    n_distinct_games = n_distinct(game_title),
    med_mean_diff_acc = med_accuracy - mean_accuracy
  ) %>%
  ungroup()

test <- last_test_assessment %>%
  left_join(test, by = "installation_id") %>%
  distinct(installation_id, game_session, game_title, .keep_all = TRUE) %>%
  pivot_wider(names_from = game_title,
              values_from = c(sum_correct, sum_incorrect, mean_accuracy, med_accuracy, sd_accuracy, n_games,
                              n_distinct_games, med_mean_diff_acc)) %>%
  select(-matches("_NA$")) %>%
  mutate_at(vars(matches("acc")),
            ~replace_na(., median(., na.rm = TRUE))) %>%
  mutate_at(vars(matches("games|sum")),
            ~replace_na(., 0)) %>%
  janitor::clean_names() %>%
  select(-c("game_session", "installation_id"))
```

## Predicting values for test set
```{r test preds}
test <- pre_recipe %>%
  bake(test)

mod_preds <- predict(rf_trn, test) %>%
  bind_cols(last_test_assessment) %>%
  select(installation_id, .pred_class) %>%
  rename(accuracy_group = .pred_class)

#write.csv(mod_preds, "submission.csv", row.names = F)
```
