---
title: "PBS Data Science EDA"
author: "EE"
date: "12/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)

#setting a function to score events that can be scored, e.g. games
score_events <- function(df){
    install_ids <- df %>% 
        distinct(game_session, installation_id, .keep_all = T) %>%
        select(game_session, installation_id, timestamp, title)
    
    df %>% 
        filter(str_detect(event_data, '""correct""')) %>%
        mutate(correct = as.numeric(str_detect(event_data, '""correct"":true'))) %>%
        select(game_session, title, event_id, event_code, correct) %>%
        group_by(game_session) %>%
        summarize(num_correct = sum(correct),
                  num_incorrect = n() - sum(correct),
                  accuracy = sum(correct) / n()) %>%
        select(game_session, num_correct, num_incorrect, accuracy) %>%
        left_join(., install_ids, by = c('game_session'))
}
```

Reading in data and cleaning
```{r data read}
train <- data.table::fread(here("2019 Data Science Bowl/Data/train.csv"))
test <- data.table::fread(here("2019 Data Science Bowl/Data/test.csv"))
trn_lbls <- data.table::fread(here("2019 Data Science Bowl/Data/train_labels.csv"))
specs <- data.table::fread(here("2019 Data Science Bowl/Data/specs.csv"))

#getting the last assessment in the test set
last_test_assessment <- test %>%
  filter(type == "Assessment") %>%
  arrange(desc(timestamp)) %>%
  distinct(installation_id, .keep_all = TRUE) %>%
  select(installation_id, title)

#filtering out the train data to only ids in the trn_lbls set
train <- train %>%
  filter(is.element(installation_id, trn_lbls$installation_id))

glimpse(train)
```

Let's see how many unique event codes there are
```{r event code explore}
train %>%
  count(event_code) %>%
  ggplot() +
  geom_col(aes(x = as.factor(event_code), y = n))
#ok, so there are some even codes that are we more frequent than others 
#I'm still not sure what event codes mean, though?
```

Let's look at different types of activities & the number of entries
```{r activity types}
train %>%
  count(type, sort = TRUE) %>%
  ggplot() +
  geom_col(aes(x = type, y = n, fill = type))
#most media types are games, followed by activities. Relatively few assessments and clips
```

But let's also see the total time spent on different activities
```{r time by activity}
train %>%
  group_by(game_session) %>%
  mutate(time_diff = game_time - lag(game_time) %>%
           replace_na(0)) %>%
  ungroup() %>%
  group_by(type) %>%
  summarize(time = sum(time_diff, na.rm = TRUE)) %>%
  ggplot() +
  geom_col(aes(x = type, y = time, fill = type))
#looks pretty similar to the above, except we see very little time spent on clips -- maybe people skip these?
```

And let's see the number of activities/types by world
--note that titles are unique. Media titles will differ by world.
```{r types by world}
train %>%
  count(world, type, sort = TRUE) %>%
  ggplot(aes(x = type, y = n, fill = type)) +
  geom_col() +
  facet_wrap(~ world)

```



Now let's look at the number of game sessions per installation id to see how often kiddos are playing
```{r game sessions by installation id}
train %>%
  distinct(installation_id, game_session) %>%
  count(installation_id, sort = TRUE) %>%
  ggplot() +
  geom_histogram(aes(n))
#big skew in the number of game sessions by installation id -- maybe we want to get rid of some of these?

#let's check out the median
train %>%
  distinct(installation_id, game_session) %>%
  count(installation_id, sort = TRUE) %>%
  summarize(median = median(n, na.rm = TRUE)) #so, the median is only 30

#let's look at what log scaling the number of game sessions does
train %>%
  distinct(installation_id, game_session) %>%
  count(installation_id, sort = TRUE) %>%
  ggplot() +
  geom_histogram(aes(log(n))) #log scaling is way better -- will want to use this
```

# Resume here

Features we can get:
-accumulated accuracy
-time spent on different types of activities
  -note that we can get the time per game session by using max(game_time) for each game_session
-accumulated number incorrect
-accumulated number correct
-(inverse) standard deviation of accuracy (?)
-log scale number of total sessions
-performance since their last assessment
