---
title: "Kaggle August 22 Tabular EDA"
description: |
    EDA of the August 22 Tabular Playground Data from Kaggle
author: "EE"
---

*add some intro text*

```{r}
library(tidyverse)
library(skimr)
library(corrplot)

theme_set(theme_minimal())

trn <- read_csv(here::here("aug_22_tabular/data/train.csv"))


```

```{r}
glimpse(trn)
```

So we have 3 character columns and the rest are numeric. The outcome, `failure`, is binary.

# Missingness


```{r}
trn |>
    summarize(across(everything(), function(x) sum(is.na(x))/length(x))) |>
    pivot_longer(
        cols = everything(),
        names_to = "var",
        values_to = "val",
    ) |>
    ggplot(aes(x = val, y = fct_reorder(var, val))) +
    geom_col() +
    scale_x_continuous(labels = scales::percent_format())
```

Ok so some missingness in here -- we'll want to explore that a bit more later.

# Outcome

```{r}
sum(trn$failure)/nrow(trn)
```

So, most outcomes are not failures -- only about 20% are failures.

# Categorical Features

```{r}
trn |>
    select(where(is.character)) |>
    summarize(across(everything(), n_distinct))
```

Ok this is really helpful -- not that many unique values for each categorical variable

```{r}
trn |>
    select(where(is.character)) |>
    pivot_longer(
        cols = everything(),
        names_to = "var",
        values_to = "val"
    ) |>
    count(var, val) |>
    ggplot(aes(x = n, y = val, fill = var)) +
    geom_col() +
    facet_wrap(vars(var), scales = "free_y") +
    theme(
        legend.position = "none"
    )
```

Alright so attribute 0 is pretty imbalanced, but product code and attribute one are more balanced.

```{r}
trn |>
    select(where(is.character), failure) |>
    mutate(failure = as.character(failure)) |>
    pivot_longer(
        cols = -failure,
        names_to = "var",
        values_to = "val"
    ) |>
    count(var, val, failure) |>
    ggplot(aes(x = n, y = val, fill = failure)) +
    geom_col() +
    facet_wrap(vars(var), scales = "free_y")
```

It's kinda hard to see what's going on here -- it might be better to compare the failure incidence in each to the baseline of ~21%

```{r}
trn |>
    select(where(is.character), failure) |>
    pivot_longer(
        cols = -failure,
        names_to = "var",
        values_to = "val"
    ) |>
    group_by(var, val) |>
    summarize(fr = sum(failure)/n()) |>
    ungroup() |>
    ggplot(aes(x = fr, y = val)) +
    geom_col(position = "dodge", fill = "steelblue") +
    geom_vline(xintercept = .21) +
    facet_wrap(vars(var), scales = "free_y")
```

Ok so even though some categories are more/less common, the failure rate across them all is pretty even.

# Numeric Features

```{r}
trn |>
    select(where(is.numeric), -id) |>
    skim()
```

ok so measurement 17 is on a much different scale from everything else. Loading is also on a strange scale

```{r}
trn |>
    select(where(is.numeric), -id) |>
    pivot_longer(
        cols = everything(),
        names_to = "var",
        values_to = "val"
    ) |>
    ggplot(aes(x = val)) +
    geom_histogram() +
    facet_wrap(vars(var), scales = "free_x")
```

Takeaways here are that attributes 2 and 3 are actually count variables, or potentially categorical variables

Loading and measurements 0 and 1 have a little bit of skew to them

Let's do the same plot again but with distributions by the outcome variable

```{r}
trn |>
    select(where(is.numeric), -id) |>
    pivot_longer(
        cols = -failure,
        names_to = "var",
        values_to = "val"
    ) |>
    mutate(failure = as.character(failure)) |>
    ggplot(aes(x = val, fill = failure)) +
    geom_histogram() +
    facet_wrap(vars(var), scales = "free_x")
```

This is a little hard to read, but I don't see anything standing out that one variable is clearly a better indicator of failure than another

```{r}
trn |>
    select(where(is.numeric), -id) |>
    cor(use = "pairwise.complete.obs") |>
    corrplot()
```

Ok so not a ton of correlations here. We see that measurement 17 has a negative correlation with some other variables, but it's possible that is because of the big scale difference?

Most importantly, failure doesn't really seem to have much of a correlation with anything

```{r}
cr_mat <- trn |>
    select(where(is.numeric), -id) |>
    cor(use = "pairwise.complete.obs")

cr_mat["failure",]
```

Yeah so very little action here...

Thoughts for modeling this data:

- Impute missing. Start with something simple like median/mode imputation, then potentially work to something else
- Dummy out categorical variables
- Normalize the continuous variables, or maybe use like the YJ transform? Some might need slight power transform.
- Possibly try PCA, but I don't know that it'll have much of an effect since the correlations here are so small
- Maybe do some feature selection via an early LASSO model?