---
title: "Disaster Tweets EDA"
author: "EE"
date: "1/6/2020"
output: html_document
---

This notebook presents an exploratory analysis of the data for the "Real or Not?" Disaster Tweets challenge

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(0408)

library(tidyverse)
library(tidytext)

train <- read_csv(here::here("Disaster Tweets/Data/train.csv"))
test <- read_csv(here::here("Disaster Tweets/Data/test.csv"))

theme_set(theme_minimal())
```

First, let's glimpse each of the datasets.
```{r trn glimpse}
glimpse(train)
```

```{r test glimpse}
glimpse(test)
```

Next, let's combine our datasets so we can explore both together
```{r}
all_df <- bind_rows(train, test) %>%
  mutate(set = if_else(!is.na(target), "tr", "te")) #the 'set' variable gives us a way to distinguish between these in our plots
```

## Missingness

Let's take a look at missingness in variables across our data
```{r missingness}
all_df %>%
  group_by(set) %>%
  summarize_all(function(x) {sum(is.na(x))/length(x)})
```
This tells us the % of missing values for each variable in each of the sets. `keyword` has a small amount of missingness in each. `location` has about 1/3rd of observations with missing values in each.

```{r missingness viz}
all_df %>%
  group_by(set) %>%
  summarize_all(function(x) {sum(is.na(x))/length(x)}) %>%
  ungroup() %>%
  pivot_longer(cols = -set) %>%
  ggplot(aes(x = name, y = value, fill = set)) +
  geom_col(position = "dodge") +
  labs(title = "% Missing by Variable")
```

## Target Frequency

Just in the train data, let's take a look at how often these tweets actually were about disasters
```{r target summary}
all_df %>%
  filter(set == "tr") %>%
  summarize(trg = mean(target))
```
About 43% of the responses are actually about disasters.

Let's also see if target values change at all with `keyword` or `location`
```{r keyword targ}
scatter_func <- function(df, group_var, n) {
  group_var <- enquo(group_var)
  
  df %>%
  filter(set == "tr") %>%
  group_by(!!group_var) %>%
  summarize(targ_freq = mean(target),
            count = n()) %>%
  arrange(desc(count)) %>%
  ungroup() %>%
  ggplot(aes(x = count, y = targ_freq)) +
  geom_point() +
  geom_text(data = . %>%
              slice(1:n),
            aes(label = !!group_var), hjust = 0, vjust = 0)
}

all_df %>%
  scatter_func(keyword, n = 30) #note that the n parameter here determines how many labels to include; not how many points to plot
```
Just looking at the top 15 keywords, we see that some seem to be pretty good at distinguishing whether a tweet describes a disaster, whereas others don't.

Let's do the same for location
```{r location targ}
all_df %>%
  filter(!is.na(location)) %>%
  scatter_func(location, n = 15)
```
Alright, so we can see that tweets in the United States (as USA or United States) seem more associated with actual disasters. Less so for cities like New York, London, or LA.

## Word Frequency

Next, we'll take a look at word frequency using some tools from the [`tidytext` package](https://CRAN.R-project.org/package=tidytext), which makes text data easier to work with using other `tidyverse` packages.

First, we'll get data to a point where we have one word per line.
```{r unnesting words}
get_words <- function(df, input) {
  input <- enquo(input)
  
  df %>%
    mutate(text = str_replace_all(!!input, "http[^[:space:]]*|@[^[:space:]]*", "")) %>% #removing links and usernames
    unnest_tokens(word, text) %>%
    anti_join(stop_words) #this will remove stop words (e.g. a, the, to) from our 'word' column
}

all_df_words <- all_df %>%
  get_words(text)
```

And now let's look at the most common words across each dataset
```{r graph common words}
all_df_words %>%
  filter(str_detect(word, "[a-zA-Z\\d:]")) %>% #filtering out weird symbols that got in
  count(set, word, sort = TRUE) %>%
  filter(n > 50) %>%
  ggplot(aes(x = word, y = n)) +
  geom_col() +
  facet_wrap(~ set, scales = "free") +
  coord_flip()
#add a sort to this
```

## RESUME HERE TOMORROW

