The goal for this analysis is to just use the text data and glmnet to predict the outcome

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(0408)

library(tidyverse)
library(tidytext)
library(glmnet)
library(tm)

train <- read_csv(here::here("Disaster Tweets/Data/train.csv"))
test <- read_csv(here::here("Disaster Tweets/Data/test.csv"))

```

Combining dfs to do some cleaning
```{r df clean}
all_df <- bind_rows(train, test)

all_df <- all_df %>%
  mutate(has_exclam = str_detect(text, "!"),
         has_quest = str_detect(text, "\\?"),
         has_hashtag = str_detect(text, "#"),
         text = str_remove_all(text, "http[^[:space:]]*"),
         text = str_remove_all(text, "@[^[:space:]]*"),
         text = str_remove_all(text, "[^[:alpha:][:space:]]*"),
         text = stripWhitespace(text),
         text = str_to_lower(text),
         text = removeWords(text, stopwords('en')),
         word_count = str_count(text, "\\w+"),
         char_count = nchar(str_remove_all(text, "[[:space:]]")),
         avg_word_length = char_count/word_count)

View(all_df)
```


Making a bag of words
```{r bag of words}
source <- VectorSource(all_df$text)
corpus <- VCorpus(source)

tweets_dtm <- as_tibble(as.matrix(DocumentTermMatrix(corpus)))
```


Now, binding back in the word bag to our original df
```{r df bind}
all_df2 <- all_df %>%
  rename(y_target = target) %>%
  select(-c("text", "keyword", "location", "id")) %>%
  bind_cols(tweets_dtm)

```

Let's get rid of very sparse columns
```{r filter sparse cols}
sparse_cols <- all_df2 %>%
  summarize_all(~mean(., na.rm = TRUE)) %>%
  select_if(. >= .002) %>%
  names() #this is a bit of a misnomer -- these are actually the non-sparse columns

all_df_use <- all_df2 %>%
  select(intersect(names(all_df2), sparse_cols))

rm(all_df2)

train_use <- all_df_use %>%
  filter(!is.na(y_target)) %>%
  select(-y_target)

train_target = train$target

test_use <- all_df_use %>%
  filter(is.na(y_target)) %>%
  select(-y_target)
```

Setting up a LASSO model
```{r lasso model}
lasso_mod <- cv.glmnet(x = as.matrix(train_use),
                       y = train_target,
                       family = "binomial",
                       alpha = 1)
```

And predicting out test values
```{r predict and write test}
test_preds <- predict(lasso_mod, as.matrix(test_use), s = "lambda.1se", type = "class")

submission <- tibble(
  id = test$id,
  target = test_preds
)

write_csv(submission, here::here("Disaster Tweets/Submissions/bag_of_words_early.csv"))
```

Note that this ends up getting a score of .8016, which is pretty good!

### Next steps

+ Handle duplicate tweets
+ Do something with location and keyword data (prob weighted mean function to post-process class value?)
+ RBERT