The goal for this analysis is to just use the text data and glmnet to predict the outcome

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(0408)

library(tidyverse)
library(tidytext)
library(glmnet)

train <- read_csv(here::here("Disaster Tweets/Data/train.csv"))
test <- read_csv(here::here("Disaster Tweets/Data/test.csv"))

```

First, let's take our train and test data and find common words.
```{r train clean}
get_words <- function(df) {
  df %>%
    mutate(text = str_replace_all(text, "http[^[:space:]]*|@[^[:space:]]*", "")) %>% #removing links and usernames
    unnest_tokens(word, text) %>%
    anti_join(stop_words)
}

test_ids <- test$id

train <- train %>%
  select(id, target, text) %>%
  get_words()

test <- test %>%
  select(id, text) %>%
  get_words()

#finding common words in both
common_words <- intersect(train$word, test$word)

trn_small <- train %>%
  filter(is.element(word, common_words))

test_small <- test %>%
  filter(is.element(word, common_words))

trn_target <- trn_small %>%
  distinct(id, target) %>%
  select(target) %>%
  as_vector()

trn_small <- trn_small %>%
  select(-target)
```

Next, let's cast our text data as a sparse matrix
```{r sparse train}
train_sparse <- trn_small %>%
  cast_sparse(id, word)

dim(train_sparse)
```

Running a glmnet model
```{r glmnet train}
trn_mod <- cv.glmnet(x = train_sparse, 
                     y = trn_target,
                     family = "binomial",
                     nfolds = 50,
                     alpha = 1)

trn_mod
```

Getting a sparse matrix of test data
```{r test sparse}
test_sparse <- test_small %>%
  cast_sparse(id, word)

dim(test_sparse)
```

Predicting test values with our model
```{r test pred}
test_preds <- predict(trn_mod, test_sparse, type = "class")
```

Binding into a csv and submitting
```{r submission}
submission <- tibble(
  id = rownames(test_preds),
  target = test_preds
) %>%
  mutate_all(as.numeric)

impute_data <- tibble(
  id = setdiff(test_ids, submission$id),
  target = 0
  )

submission <- submission %>%
  bind_rows(impute_data) %>%
  arrange(id)

write_csv(submission, here::here("Disaster Tweets/Submissions/glmnet_initial.csv"))
```

Note that this actually doesn't end up being a very good model :)