---
title: "Cat Encoding 2 EDA"
author: "EE"
date: "2/7/2020"
output: html_document
---

```{r initial}

library(tidyverse)
library(lubridate)
library(vroom)

train <- vroom(here::here("Categorical Feature Challenge 2/Data/train.csv"))

test <- vroom(here::here("Categorical Feature Challenge 2/Data/test.csv"))

miss_plot <- function(data, color1 = "steelblue1", color2 = "steelblue4", bound = 0) {
  miss_tab <<- tibble(
    column = names(data),
    perc_miss = map_dbl(data, function(x) sum(is.na(x))/length(x))
  ) %>%
    filter(perc_miss > bound)
  
  ggplot(miss_tab, aes(x = column, y = perc_miss)) +
    geom_bar(stat = "identity", aes(fill = ..y..)) +
    scale_y_continuous(labels = scales::percent) +
    theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
    scale_fill_gradient(low = color1, high = color2, name = "Percent Missing") +
    labs(
      title = "Missingness by Variable",
      y = "Percent Missing",
      x = "Variables"
    )
}
```

The point of this kernel is to do an EDA of the data to inform how we build our models later on.

### Cleaning

Doing a bit of transforming and cleaning
```{r reshaping}
targ <- train$target

train <- train %>%
  mutate(target = as.factor(target)) %>%
  mutate(bin_4 = if_else(bin_4 == "Y", TRUE, FALSE)) %>%
  mutate_at(vars(matches("bin")), ~as.logical(.)) %>%
  mutate_at(vars(matches("nom|ord")), ~as_factor(.))

```


### Target Mean

First, let's look at the distribution of the outcome on its own.
```{r}
mean(targ)

ggplot(train) +
  geom_bar(aes(target, fill = target))
```

So, about a fifth of all responses are "hits" for the target.

### Unique Counts

Next, let's take a look at how many unique values we have for each variable
```{r distinct counts}
train %>%
  summarize_all(n_distinct)
```

Alright, so, including our `NA` values, we have anywhere from 3 distinct values per column (for the binary columns) to 2219 distinct values per column (for our nom_9 variable). One interesting thing to note here is that "day" only has 8 unique values, which is fewer than I would have guessed.

Another thing potentially worth noting is that ordinal variables 3, 4, and 5 have 16, 27, and 191 unique values, respectively, and so we can probably treat those as continuous. We could possibly also treat ord_2, which has 7 distinct values, as continuous as well.

### Missingness

We'll probably want to see how much missingness is in each variable. The `miss_plot()` function defined above will give us a plot with the proportion of missingness for each variable in our dataset
```{r missingness}
miss_plot(train)
```

Ok, so, pretty uniformly we have around 3% missing, and this looks like it holds for all of our features. This shouldn't be too hard to deal with.

### Binary Variables

We can check out the distribution of our binary variables now, which will be the easiest to visualize. We'll start with each variable independently, then look at how they relate to the outcome, and then examine the extent to which they're correlated

```{r binary univariate}
train %>%
  select(matches("bin")) %>%
  pivot_longer(cols = everything(),
               names_to = "key",
               values_to = "vals") %>%
  count(key, vals) %>%
  ggplot() +
    geom_col(aes(fill = vals, x = key, y = n), position = "dodge")
```

```{r}
train %>%
  select(matches("bin|targ")) %>%
  pivot_longer(cols = matches("bin"),
               names_to = "key",
               values_to = "vals") %>%
  count(key, target, vals) %>%
  ggplot() +
  geom_col(aes(fill = vals, x = key, y = n), position = "dodge") +
  facet_wrap(~target)
  
```

```{r}
train %>%
  select(matches("bin")) %>%
  mutate_all(as.numeric) %>%
  cor(use = "pairwise.complete.obs")
```

Ok, so these are pretty much uncorrelated

### Ordinal Variables

Let's take a look at our ordinal variables now. I also noticed that I was writing the same code a lot earlier, so I'll write a function to select, pivot longer, and count

```{r}
sel_piv_cnt <- function(df, vars) {
  df %>%
  select(matches(vars)) %>%
  pivot_longer(cols = matches(vars),
               names_to = "key",
               values_to = "vals") %>%
  count(key, vals)
}
```

Let's visualize counts of our ordinal variables now

```{r}
train %>%
  sel_piv_cnt("ord") %>%
  ggplot() +
    geom_col(aes(fill = vals, x = vals, y = n), position = "dodge") +
  facet_wrap(~key, scales = "free") +
  theme(
    legend.position = "none"
  )

```

And let's also see the distribution of the target variable
```{r}
train %>%
  select(matches("ord|targ")) %>%
  pivot_longer(cols = matches("ord"),
               names_to = "key",
               values_to = "vals") %>%
  count(key, target, vals) %>%
  ggplot() +
  geom_col(aes(fill = target, x = vals, y = n), position = "dodge") +
  facet_wrap(~key, scales = "free") +
  theme(
    legend.position = "none"
  )
```

### Nominal Variables

```{r}
train %>%
  sel_piv_cnt("nom") %>%
  ggplot() +
    geom_col(aes(fill = vals, x = vals, y = n), position = "dodge") +
  facet_wrap(~key, scales = "free") +
  theme(
    legend.position = "none"
  )
```

```{r}
train %>%
  select(matches("nom|targ")) %>%
  pivot_longer(cols = matches("nom"),
               names_to = "key",
               values_to = "vals") %>%
  count(key, target, vals) %>%
  ggplot() +
  geom_col(aes(fill = target, x = vals, y = n), position = "dodge") +
  facet_wrap(~key, scales = "free", shrink = TRUE) +
  theme(
    legend.position = "none"
  )
```

This is still a work in progress -- I'll plan to pick this up later to further investigate the data.